Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-31 transmission.tex
include kappa time dependencies

The time-varying component, $\beta(t)$, allows for behavioural changes that impact transmission but are not captured by the contact matrices (\eg mask wearing).
I impose a random walk prior on $\log\beta(t)$ on the weekly timescale, smoothing its value over time.
This prior encodes that behaviour changes smoothly over time.
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk (see \cref{SEIR:sec:priors}).
$\beta(t)$ is the only time-varying fitted parameter.
Therefore, in practice, it will also absorb other effects such as any error in $\matr{\kappa}$.

@@

The only component of $\matr{\beta}$ that allows age-dependent heterogeneity in the temporal changes is $\matr{\kappa}$.
It is plausible that this would not capture all behavioural changes.
For example, the introduction and relaxation of control measures within schools.
$\matr{\kappa}$ only captures changes in in-school transmission through changes in school attendance, not incorporating the probability of transmission, which could be modelled as changes in $q_{aa}$ for $a$ of school age.
Similarly, there is evidence that $q_c$ is different for Alpha and pre-Alpha variants~\autocite{zhuRole}, which could again motivate these changing over time.

>>>>

The time-varying component, $\beta(t)$, allows for behavioural changes that impact transmission but are not captured by the contact matrices (\eg mask wearing).
I impose a random walk prior on $\log\beta(t)$ on the weekly timescale, smoothing its value over time.
This prior encodes that behaviour changes smoothly over time.
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk (see \cref{SEIR:sec:priors}).
$\beta(t)$ is the only time-varying fitted parameter.
Therefore, in practice, it will also absorb other effects such as any error in $\matr{\kappa}(t)$.

@@

The only component of $\matr{\beta}$ that allows age-dependent heterogeneity in the temporal changes is $\matr{\kappa}(t)$.
It is plausible that this would not capture all behavioural changes.
For example, the introduction and relaxation of control measures within schools.
$\matr{\kappa}(t)$ only captures changes in in-school transmission through changes in school attendance, not incorporating the probability of transmission, which could be modelled as changes in $q_{aa}$ for $a$ of school age.
Similarly, there is evidence that $q_c$ is different for Alpha and pre-Alpha variants~\autocite{zhuRole}, which could again motivate these changing over time.
